---
title: "Problem Sheet 3"
subtitle: "Theory of Inference"
author: "Dom Hutchinson"
header-includes:
   - \usepackage{dsfont}
   - \usepackage{fancyhdr}
   - \pagestyle{fancy}
   - \renewcommand{\headrulewidth}{0pt}
   - \fancyhead[L]{Dom Hutchinson}
   - \fancyhead[C]{Theory of Inference - Problem Sheet 3}
   - \fancyhead[R]{\today}
   - \newcommand{\dotprod}[0]{\boldsymbol{\cdot}}
   - \newcommand{\cosech}[0]{\mathrm{cosech}\ }
   - \newcommand{\cosec}[0]{\mathrm{cosec}\ }
   - \newcommand{\sech}[0]{\mathrm{sech}\ }
   - \newcommand{\prob}[0]{\mathbb{P}}
   - \newcommand{\nats}[0]{\mathbb{N}}
   - \newcommand{\cov}[0]{\mathrm{Cov}}
   - \newcommand{\var}[0]{\mathrm{Var}}
   - \newcommand{\expect}[0]{\mathbb{E}}
   - \newcommand{\reals}[0]{\mathbb{R}}
   - \newcommand{\integers}[0]{\mathbb{Z}}
   - \newcommand{\indicator}[0]{\mathds{1}}
   - \newcommand{\nb}[0]{\textit{N.B.} }
   - \newcommand{\ie}[0]{\textit{i.e.} }
   - \newcommand{\eg}[0]{\textit{e.g.} }
   - \newcommand{\X}[0]{\textbf{X}}
   - \newcommand{\x}[0]{\textbf{x}}
   - \newcommand{\iid}[0]{\overset{\text{iid}}{\sim}}
   - \newcommand{\proved}[0]{$\hfill\square$\\}
output:
  pdf_document: 
     fig_width: 6
     fig_height: 4
  html_notebook: default
  html_document: default
  word_document: default
---
# Question 1

## Part a)
\[\begin{array}{rrrl}
&\|\tilde{\textbf{y}}-\tilde{\textbf{X}}\beta\|^2&=&\sum_i(\tilde{y}_i-\sum_j\tilde{X}_{ij}\beta_j)^2\\
&&=&\sum_i(W_{ii}y_i-\sum_jW_{ii}X_{ij}\beta_j)^2\\
&&=&\sum_iW_{ii}^2(y_i-\sum_jX_{ij}\beta_j)^2\\
\text{Define}&W_{ij}&:=&\begin{cases}\frac1{\sqrt{n_{se(i)}}}&i\equiv j\\0&\text{otherwise}\end{cases}\\
\implies&\|\tilde{\textbf{y}}-\tilde{\textbf{X}}\beta\|^2&=&\sum_i\frac1{\sqrt{n_{se(i)}}}(y_i-\sum_j X_{ij}\beta_j)^2
\end{array}\]

## Part b)
\[\begin{array}{rrcl}
\text{We have }&\|\tilde{\textbf{y}}-\tilde{X}\beta\|^2&=&(\tilde{\textbf{y}}-\tilde{X}\beta)^T(\tilde{\textbf{y}}-\tilde{X}\beta)\\
&&=&(y-X\beta)^TW^2(y-X\beta)\\
&&=&y^TW^2y-y^TW^2X\beta-\beta^TX^TW^2y+\beta^TX^TW^2X\beta\\
\implies&\frac\partial{\partial\beta}\|\tilde{\textbf{y}}-\tilde{X}\beta\|&=&-X^TW^2y-X^TW^2y+2X^TW^2X\beta\\
\text{Setting}&0&=&\frac\partial{\partial\beta}\|\tilde{\textbf{y}}-\tilde{X}\beta\|\\
\implies&0&=&2(X^TW^2X\hat\beta_\text{WLS}+X^TW^2X\hat{\beta}_\text{WLS})\\
\implies&\hat{\beta}_\text{WLS}&=&(X^TW^2X)^{-1}X^TW^2y\\
\implies&\expect[\hat{\beta}_\text{WLS}]&=&\expect((X^TW^2X)^{-1}X^TW^2y)\\
&&=&(X^TW^2X)^{-1}X^TW^2\expect(y)\\
&&=&(X^TW^2X)^{-1}X^TW^2X\beta\\
&&=&\beta
\end{array}\]
Since $\expect(\hat\beta_\text{WLS})=\beta$ then $\hat\beta_\text{WLS}$ is an unbiased estimator of $\beta$.

## Part c)
Since this estimator is unbiased and fulfils all the residual assumptions I would recommend the weighted least squares estimator here as it takes into account more information than a standard least squares estimator.

# Question 2
Model doesn't account for the fact that the markets close at night. There are likely to be sizable changes in the price of a share between closing and opening of the markets. This could be seen by regular spikes in model residual values at the start of each day.

# Question 3

## Part a)
If we know that an observation is not in group one or two, then it most be in group three. Thus the groups are not independent \& the model matrix is rank deficient. This means we cannot invert the matrix.

## Part b)
```{r}
m1<-lm(weight ~ group,data=PlantGrowth)
```
In this interpretation value which is the expected value for observations in $\mathtt{ctrl}$ group. There are then two binary variables which state whether the observation was made from $\mathtt{grouptrt1}$ or $\mathtt{grouptrt2}$ (these are not independent). The parameters associated with each of these variables is the expected difference in $\mathtt{weight}$ for observations in that group, compared to $\mathtt{ctrl}$.

## Part c)
$H_0:\pmb\beta=\begin{pmatrix}\mu\\0\\0\end{pmatrix}$ against $H_1:\pmb\beta\neq\begin{pmatrix}\mu\\0\\0\end{pmatrix}$.
```{r}
anova(m1)
```
We reject null hypothesis at the 5% level. Thus concludinf that $\mathtt{group}$ is needed in the model.\
It uses the results in second $3.4.4$ of the notes, in particular
$$F=\dfrac{(RSS_0-RSS_1)/q}{RSS_1/(n-p)}$$
where $n=30,\ p=3,\ q=3$. $TODO check this

## Part d)
```{r}
m0<-lm(weight~1,data=PlantGrowth)
anova(m0,m1)
```
Same result as in part c).

## Part e)
```{r}
summary(m1)
```
$H_0:\pmb\beta=\pmb0$ against $H_1:\pmb\beta\neq\pmb0$. The $p$-Value is unchanged.

## Part f)
We have that
$$\dfrac{\hat\beta_2-\beta_i}{\hat\sigma_{\hat\beta_2}}\sim t_{27}$$
Thus
\[\begin{array}{rcl}
1-2\alpha&=&\prob\left(-t_{27}(\alpha)<\dfrac{\hat\beta_2-\beta_i}{\hat\sigma_{\hat\beta_2}}<t_{27}(\alpha)\right)\\
&=&\prob(\hat\beta_2-t_{27}(\alpha)\hat\sigma_{\hat\beta_2}<\beta_2<\hat\beta_2+t_{27}(\alpha)\hat\sigma_{\hat\beta_2})
\end{array}\]
In this scenario we have that
\[\begin{array}{rcl}
\hat\beta_2&=&\bar{x}_\text{trt2}-\bar{x}_\text{ctrl}\\
&=&5.525-5.032\\
&=&0.4940\\
\hat{\sigma}^2_{\hat\beta_2}&=&\frac1{10-1}\sum_{i=1}^{10}(y_i-\hat{y}_i)^2\\
&=&\frac{1}{9}1.76284\\
&=&0.1959\\
\implies\hat\sigma_{\hat\beta_2}&=&0.4425733
\end{array}\]
Giving us the following $90\%$ confidence interval for difference between treatment 2 and the control
\[\begin{array}{rl}
[0.4940-1.703\cdot0.4425733,0.4940+1.703\cdot0.4425733]\\
=&[-0.272,4.342]
\end{array}\]